---
title       : Developing Data Project
subtitle    : Mcars Prediction  
author      : Abdul
job         : Student
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---
1. Summary

There is a set of variables related to miles per gallon (MPG) (outcome).

Here the difference of the MPG between automatic and manual transmissions is focused, the mtcars dataset in car package is taken as the data source, and techniques about regression models is used to solve the following two questions:

   (a) "whether an automatic or manual transmission better for MPG"
   (b) "Quantifying the different between the MPG automatic and manual transmissions?"

2 Part 2 of The Project
2.1 Load and Test Data

    The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

```{r}  
library(datasets)
mpgData <- with(mtcars, data.frame(mpg, am))
mpgData$am <- factor(mpgData$am, labels = c("Automatic", "Manual"))
#summary(mpgData)
```

2.2 Data Processing 
2.2.1 (a): whether an automatic or manual transmission better for MPG?

```{r}
summary(mpgData[mpgData$am == "Automatic",])
```
```{r}
summary(mpgData[mpgData$am == "Manual",])
```

2.2.2 (b): Quantifying the different between the MPG automatic and manual transmissions?

```{r}
fit <- lm(mpg ~ as.integer(am), data=mpgData)
summary(fit)
```

---
3 Analysis

It appears that the manual transmission is better than automatic MPG, by of increased by 7.2449.
4 Appendix
3.1 Fig 1. MPG between automatic and manual transmissions

```{r}
par(mfrow = c(1, 2))
with(mtcars, boxplot(mpg/30, disp/80, hp/80, drat, wt, qsec/20, main = "Scaled boxplots; mtcars"), 
    xlab = "Continuous variable", ylab = "Scale (adjusted to allow simultaneous display")
axisnames <- c("MPG/10", "DISP/50", "HP/50", "drat", "wt", "qsec/5")
axis(1, at = c(2, 4, 6, 8, 10, 12), labels = axisnames, las = 2.5, cex.axis = 1.0, 
    tck = 0.05)
pairs(mtcars)
```

From scaled boxplots that the continuous variables are skew, and outlying values for horsepower, weight, and quarter mile speed. Transformations could be required. lets note from the Appendix that  categorical variables have sufficient frequencies in all cells for  regression models and are ordinal, so usable in linear models.  
Also lets examine bivariate plots of variables.The relationship between (automatic vs manual) transmission and fuel consumption ; manual transmissions yielding higher values of MPG.  However, there are  relationships between fuel consumption and other variables (cylinders, engine displacement, weight).  Transmission  looks to be correlated with some of these potential predictors, an important question is to what extent transmission independently affects fuel economy. 
Now ascertain whether fuel consumption differs between manual and automatic transmission cars, and quantify this.

---
Fig 2.
```{r}
par(mfrow=c(1,2))
with(mpgData,{
     boxplot(mpg ~ am, 
             ylab = "miles per gallon (MPG)")
     plot(mpg ~ as.integer(am),
             xlab = "Automatic (1) or Manual(2)",
             ylab = "miles per gallon (MPG)")
     abline(fit, col=2)
     })
```

Fig 3.
```{r}
mtcarsregmodel <- lm(mpg ~ am, data = mtcars)
summary(mtcarsregmodel)
```

Transmission accounts for 36% of the variance in fuel consumption, that's statistically significantly better than chance (p < .0003). The intercept and slope coefficients shows that automatic transmissions achieve 17.147 miles per gallon fuel economy and that manual transmission cars achieve 24.39 miles per gallon fuel economy, on average.
These point estimates, are expressed at a 95% Confidence Intervals, for which we use the t.test method in R

---
4 Appendix
4.1 
```{r}
with(mtcars, table(cyl))
```

```{r}
with(mtcars, table(am))
```

```{r}
with(mtcars, table(gear))
```

```{r}
with(mtcars, table(carb))
```

```{r}
with(mtcars, table(vs))
```

